---
title: <small>An introduction to digital open <br>research in clinical psychology</small>
author: <small>Matthias Weiler</small>
date: <small>01.04.2019</small>
output:
  revealjs::revealjs_presentation:
    theme: black
    center: TRUE
    self_contained: FALSE
    incremental: TRUE
    reveal_plugins: ["chalkboard"]
    css: styles.css
    reveal_options:
      chalkboard: 
      readonly: FALSE
---


# <small> What did we do last session?</small>

----

We looked at ways to  
analyze your data

----

... created more R Scripts

----

... and connected them to R Notebooks

----

Any questions or anything to add?



# <small>Last session's homework</small>

----

Who of you created an R Notebook  
to document your code?

----

Who of the team captains added  
their team members as collaborators  
to their GitHub repository?

----

.. added their team members as  
contributors to their research  
projects in OSF?

----

Any questions regarding  
last session's homework?



# <small>Today:<br>Data visualization</small>



# <small>Graphics with ggplot2</small>



# <small>Four steps to visualization</small>

----

## <small>1. Call ggplot()</small>

```{r, echo = FALSE}
library(ggplot2)
```

```{r}
ggplot()
```

----

## <small>2. Define data</small>

```{r}
ggplot(data = mpg)
```

----

## <small>3. Define aesthetic mappings</small>

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy))
```

----

## <small>4. Add geometric object(s)</small>

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_point()
```


# <small>Some examples</small>

----

## <small>Histograms</small>

```{r message=FALSE}
ggplot(data = mpg, aes(x = hwy)) + 
  geom_histogram()
```

----

## <small>Frequency polygons</small>

```{r message=FALSE}
ggplot(data = mpg, aes(x = hwy)) + 
  geom_freqpoly()
```

----

## <small>Densityplots</small>

```{r}
ggplot(data = mpg, aes(x = hwy)) + 
  geom_density()
```

----

## <small>Scatterplots</small>

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_point()
```

----

## <small>Lineplots</small>

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_line()
```

----

## <small>Maps</small>

```{r}
ggplot(map_data("world"), aes(long, lat, group = group)) +
  geom_polygon() 
```

----

## <small>Dinosaurs</small>

```{r, echo=FALSE, message=FALSE}
library(datasauRus)
library(dplyr)
```

```{r, message=FALSE}
dino <- datasaurus_dozen %>% filter(dataset == "dino")
ggplot(dino, aes(x = x, y = y)) +
  geom_point(color = "red") +
  theme_void()
```


# <small>More examples</small>

----

## <small>Scatterplots</small>

```{r}
ggplot(data = mpg, aes(x = drv, y = hwy)) + 
  geom_point()
```

----

## <small>Jitterplots</small>

```{r}
ggplot(data = mpg, aes(x = drv, y = hwy)) + 
  geom_jitter()
```

----

## <small>Boxplots</small>

```{r}
ggplot(data = mpg, aes(x = drv, y = hwy)) + 
  geom_boxplot()
```

----

## <small>Violinplots</small>

```{r}
ggplot(data = mpg, aes(x = drv, y = hwy)) + 
  geom_violin()
```

----

## <small>Beeswarmplots</small>

```{r}
ggplot(data = mpg, aes(x = drv, y = hwy)) + 
  ggbeeswarm::geom_beeswarm()
```



# <small>Combining plots</small>

----

## <small>Histogramm</small>

```{r, message=FALSE}
ggplot(data = mpg, aes(x = hwy)) + 
  geom_histogram()
```

----

## <small>Add a frequency polygon</small>

```{r, message=FALSE}
ggplot(data = mpg, aes(x = hwy)) + 
  geom_histogram() +
  geom_freqpoly()
```

----

## <small>Scatterplot</small>

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_point()
```

----

## <small>Add a regression line</small>

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

----

## <small>Add a second regression line and conficence interval</small>

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(method = "loess", se = TRUE)
```

## <small>Violin plot</small>

```{r}
ggplot(data = mpg, aes(x = drv, y = hwy)) + 
  geom_violin()
```

----

## <small>Add a box plot</small>

```{r}
ggplot(data = mpg, aes(x = drv, y = hwy)) + 
  geom_violin() +
  geom_boxplot(width = 0.1)
```



# <small>Edit your axis</small>

----

## <small>Axis limits</small>

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_point() +
  scale_x_continuous(limits = c(0, 10))
```

----

## <small>Axis labels</small>

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_point() +
  labs(x = "Engine displacement", y = "Highway miles per gallon")
```

----

## <small>Axis scales</small>

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_point() +
  scale_x_continuous(limits = c(0, 10), breaks = c(0:10))
```



# <small>Other labels</small>

----

## <small>Title</small>

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
  geom_point() +
  labs(title = "My Title")
```

----

## <small>Legend</small>

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy, color = class)) + 
  geom_point() +
  labs(color = "My Legend")
```



# <small>Any questions?</small>



# <small>Preview: next session</small>

----

"Putting it all together"



# <small>Homework</small>

----

Add graphics to your slides

----

Add graphics to your R Notebooks



# <small>Any questions?</small>



# <small>Thank you for your attention</small>
